The Sloan Digital Sky Survey Spectroscopic Reduction Pipeline

### 2005-02-05 doc/paper/skysubtraction.tex
filled in some extraction text

@@

The accuracy of the SDSS spectral sky subtraction has been tested in recent
scientific analyses using large numbers of background-limited spectra.   
The distribution of flux residuals between spectral observations 
on separate nights has been shown to be 6\%-9\% larger in width than expected
from the reported errors (McDonald et al. 2004, Burgess \& Burles 2005).
For combined observations of faint galaxies, the distribution of residuals 
about the best model fit is well described by a gaussian with exponential
tails (Bolton et al. 2004), where the magnitude of tails is greatly enhanced in
the vicinity of strong sky emission lines.

>>>>

\subsection{Extraction}

We utilize an optimal extraction procedure to transform a 2-dimensional image of
the fiber spectrograms to a set of 320 1-dimensional spectra.  We use the same procedure
on all processed frames, and will describe the details in processing the flat-field frames, the
arc lamp frames, and finally the science frames.  
The extraction is performed row-by-row, so each frame requires an outside loop over 2048 individual rows.
The counts in each row of 2048 pixels are modeled with a linear combination of 320 profiles plus a low-order
polynomial background.  We decided to perform profile fitting to the SDSS spectral images 
for the following reasons:

\item{The fiber profiles overlap significantly with their nearest neighbors.}

\item{The profiles are relatively stable and smoothly vary over location on the CCDs.  
They result from the convolution of the fiber illumination pattern + the point-spread function
of the optics in the SDSS spectrographs + the window funciton of the Site CCD 24 micron square pixels. }

\item{A least-squares fit with the true profile delivers an unbiased and minimum variance estimate of the counts
(e.g. Horne et al. 1995).}

The major obstacle to actual optimal extraction is the determination of the exact shape and position 
of the underlying profile.  We investigated a range of symmetric, analytic profiles with a 
minimum number of free parameters.  In the end, we found the best results using a simple gaussian 
or a slightly steeper profile characterized by $e^{-|x|^3/\sigma^3}$.  The profiles
are defined to be normalized to unity, so the amplitude of the profile fit gives a direct estimator
of the counts at a single spectral position.

Extraction is first carried out on all flat-field exposures (typically 10 second observations of
4 quartz lamps projected on petals covering the entrance aperture of the telescope) which were observed
with a common plug plate configuration.  The flat-field extraction is used to determine the shape of
the profile as a function of wavelength and position on the CCD.  The resulting spectral extractions are also used to 
determine the empirical fiber-to-fiber variations in throughput (as a function of wavelength) 
during a single setup.

Three types of exposures, corresponding to flat-field, arc lamp and science frames,
 are extracted with the same basic procedure
which we describe below.  The frames associated with a single observing sequence 
are processed in a certain order.  The first types to be extracted are the 
flat-fields frames ("FLATS").

@@

The accuracy of the SDSS spectral sky subtraction has been tested in recent
scientific analyses using large numbers of background-limited spectra.   
The distribution of flux residuals between spectral observations 
on separate nights has been shown to be 8\% larger in width than the average expected
from the reported errors (McDonald et al. 2004, Burgess \& Burles 2005).
The spectra of faint, background-limited galaxies observed with SDSS exhibit a 
distribution of residuals about the best fit spectral models described by the sum of a
Gaussian and Laplace distribution (Bolton et al. 2004).  
The influence of the exponential tail is greatly enhanced in the vicinity of strong sky 
emission lines representative of systematic errors in the brightest lines, but the gaussian 
width closely is closely represented by the reported variance array.
